---
title: Saving BumpyMatrices to file
author:
- name: Aaron Lun
  email: infinite.monkeys.with.keyboards@gmail.com
package: alabaster.bumpy 
date: "Revised: January 1, 2024"
output:
  BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Saving and loading BumpyMatrices}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE}
knitr::opts_chunk$set(error=FALSE, warning=FALSE, message=FALSE)
library(BiocStyle)
self <- Githubpkg("ArtifactDB/alabaster.bumpy")
```

# Overview

The `BumpyMatrix` class provides a representation of complex ragged data structures - see the `r Biocpkg("BumpyMatrix")` package for more information.
This is used to coerce immune repertoire, spatial transcriptomics and drug response data into a familiar 2D array for easy manipulation.
The `r self` package allows users to save a `BumpyMatrix` to file within the [**alabaster** framework](https://github.com/ArtifactDB/alabaster.base).

# Saving a `BumpyMatrix`

Let's make a `BumpyMatrix` to demonstrate:

```{r}
library(BumpyMatrix)
library(S4Vectors)
df <- DataFrame(x=runif(100), y=runif(100))
f <- factor(sample(letters[1:20], nrow(df), replace=TRUE), letters[1:20])
mat <- BumpyMatrix(split(df, f), c(5, 4))
```

Saving it to file involves calling `saveObject`:

```{r}
library(alabaster.bumpy)
tmp <- tempfile()
saveObject(mat, tmp)
list.files(tmp, recursive=TRUE)
```

# Loading a `BumpyMatrix`

The loading procedure is even simpler as the metadata of the saved `BumpyMatrix` remembers how it was saved.
We can just use `alabaster.base::readObject()` or related functions, and the R interface will automatically do the rest.

```{r}
readObject(tmp)
```

# Session info {-}

```{r}
sessionInfo()
```
